Rutgers Autonomous Aircraft Team\\Technical Report\\2010 AUVSI UAS Competition

### 2010-05-25 paper/paper.tex
changed pantilt dimensions in text based on new measurments, new figure 6b
cogans changes

\documentclass[10pt,twocolumns]{report}
\usepackage[pdftex]{graphicx}
\usepackage[left=1.5in, right=1.5in, top=1.25in, bottom=1.25in]{geometry}
\usepackage{hyperref}
\hypersetup{ colorlinks=true
           , linkcolor=blue
           , pdfnewwindow=true 
           }
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{float}
\usepackage{textcomp}
\usepackage{multicol}
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
\newcommand{\itemspace}{	\setlength{\itemsep}{0cm} \setlength{\parskip}{0cm}}
\newcommand{\degrees}[1]
{
\begin{math}
#1^{\circ} 
\end{math}
}
\hyphenation{auto-nomous construction}

@@

We are advised by
 Professor Tobias Rossmann,
 and several members of the Tri-County R/C Club \cite{tricountyRC}:
 Larry Kosinar,
 Bill Perkowski,
and  Mike Korb, 
%and Chris (test pilot) 
amongst many who spent time mentoring us.

@@

We would like to acknowledge
Louis Stumpf, for donating the aircraft which became the \emph{Daedalus};
Jeff Perkins, for his prolific networking and promotion of the team;
Gregg Peters of BP Hobbies for getting us out of numerous binds;
Dave Devries and Michael Maia, of Invensense, for their technical guidance;
Professor Chung-chieh Shan, for overseeing Pat's independent study on the IMU design;
Ivan Seskar, for keeping us (and many other undergrads) in the black;
Roger Kondos and Brij Pathak, for contributing to the thermopile attitude sensor.

\begin{figure}[b!]
	\centering
	\includegraphics[width=0.6\textwidth]{../images/wreckage.jpg}
	\caption{The recovery of our 2009 entry, the \emph{Icarus}}
	\label{fig:wreckage}
\end{figure}

\newpage
\section{Systems Engineering}
\subsection{Goals}
For this entry, we took a close look at our goals and requirements, and dialed back our approach to ensure we would meet the most basic competition requirements successfully.
We started the design by forming, as a team, some broad goals based on our experience in the 2009 competition.
These formed the basis on which sub-teams and individuals tackled the smaller problems while still keeping an eye on the big picture.

We also wanted to balance building with buying, keeping in mind our slim finances. While we couldn't build every part of our entry from scratch, we tried to build whatever we could given the 16th June competition deadline.

%\footnote{``If you want to make an apple pie from scratch, you must first create the universe'' - Dr. Carl Sagan \cite{saganquote}},

@@

Our team was divided by area of expertise into a mechanical team and an electrical team.The mechanical team began the year with the goal of producing two competition worthy airframes. These were to become the \emph{Daedalus} and \emph{Knight Two}.

The teams made design decisions based on the aforementioned competition goals.

One of the most difficult design decisions was to quantify ``ample payload.'' Payload size and weight is the major determining factor in aircraft size.  We originally assumed a 2.5lb electronics (computer + batteries) payload and a 2.5lb pan-tilt camera 8'' in diameter, requiring 6'' of vertical space extruding from the bottom of the aircraft. These considerations introduced two major design parameters: the aircraft must be capable of carrying a minimum 5lb payload, and must be situated to provide clearance for the pan tilt unit.

Additionally, we sought an easily controllable, stable platform to accommodate the autopilot, and an ability to fly at low speeds, in order to more easily capture clear images.

We determined, after consulting with R/C experts, that a high wing trainer is the most stable and proven fixed-wing design. Trainers traditionally have light wing loading for slow speed flight, and are able to carry a comparatively large payload. Additionally, a large plane with lower wing loading is easier to control, and easier to recover if power is lost. Consciously agreeing to err on the side of caution, we sought a large plane capable of accommodating a worst-case scenario payload, and settled on a wingspan on the order of 10'-12'. Though lighter and smaller payload alternatives were found as the year progressed, these merely leave us with more freedom for future modifications.
%That last sentence sounds kind of eh, I'm stuck on a decent wording, so I'm leaving it for now.

The decision to employ a large, high wing trainer, however, forced us to rethink our original pan-tilt design. Last year, in order to accommodate the pan-tilt unit, we designed a pusher plane, enabling us to mount the pan-tilt under the plane's nose, which had sufficient clearance. Since trainers are generally designed with front-mounted engines (I presume?), we were forced to engineer a new solution requiring significantly less clearance. In the long run, we modified the pan-tilt mount to allow us to recess more of the pan-tilt unit into the body of the plane---this had the added benefit of providing more protection to an expensive piece of equipment on takeoff and landing.

@@

We felt the Paparazzi system was full featured and proven, but we weren't satisified with the capabilities of the existing Paparazzi hardware. The ``Tiny'' v2.1\cite{paparazzi_tinyv21}, the latest

@@

\subsubsection{Daedalus}
\begin{wrapfigure}{l}{0.4\textwidth}
	\includegraphics[width=0.37\textwidth]{../images/daedalus_isometric.jpg}
	\caption{The \emph{Daedalus}}
	\label{fig:daedalus}
\end{wrapfigure}
A member of our local R/C club, Tri-County RC \cite{tricountyRC}, donated a ten-foot wingspan high wing trainer to our club. The plane, first built out of foam and balsa from long-lost plans in 1986, was accepted graciously, but required quite a bit of work before it was once again flight-worthy. We removed the covering to find water damage and rot which resulted from years of storage. We stripped and rebuilt nearly the entire airframe, adding carbon-fiber reinforcements to the wing in anticipation of increased wing loading due to our payload. Once rebuilt and covered, we tested the \emph{Daedalus} weighed 19.5 lbs with engine. A full 24oz fuel tank, plus 3.5 lbs electronics payload, would bring the maximum takeoff weight to 25.5lbs.

@@

 \begin{wrapfigure}{r}{0.4\textwidth}
	\includegraphics[width=0.37\textwidth]{../images/daedalus_sandbagtest.jpg}
	\caption{Load testing the underside of the \emph{Daedalus}'s wings. Shown with 16lbs of load.}
	\label{fig:sandbagtest}
\end{wrapfigure}
The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.  The \emph{Daedalus} wings supported 64 lbs (250\% of the maximum take-off weight) under static loading.  We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight. Our post-test structure inspection found no warped members or cracked joints.

@@

\begin{wrapfigure}{l}{0.4\textwidth}
	\centering
	\subfloat[The unfinished \emph{Knight Two}]{\label{fig:knighttwo_construction} 
			\includegraphics[width=0.37\textwidth]{../images/knighttwo_uncovered_outdoors.jpg}}
\\
	\subfloat[A completed 12 Foot Telemaster. Photo Credit: Hobby Lobby] {\label{fig:knighttwo_12ft_telemaster} 
			\includegraphics[width=0.37\textwidth]{../images/telemaster_12ft_hobbylobby.jpg}}
  	\label{fig:knighttwo}
  	\caption[Knight Two]{Knight Two}

\end{wrapfigure}
Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations

@@

@@

We took advantage of the MHS-PM5's swiveling lens assembly 
(see Figure \ref{fig:camera_frontview}, and added a servo actuator which could rotate the lens \degrees{135} back from straight forward, as shown in Figure \ref{fig:pantilt_linkdetail}. The tilt servo is a Hitech HS 77BB \ref{hitechs77bb}, and is capable of \degrees{180} rotation. Figures \ref{fig:pantilt} and \ref{fig:pantilt_linkdetail} show the camera lens tilted at \degrees{45} below the horizon.
The camera and tilt servo assembley is attached to the aircraft on a rotating mount. A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembly. The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{1260} rotation (3.5 revolutions). Both servos are powered and controlled by the Pan-tilt control board.

@@

\subsection{Safety Observers}
A safety observer is defined as a person, devoid of any additional specific responsibilities, who remains in the starting area during all pre-flight procedures.  The safety observer is expected to watch for unsafe conditions, included but not limited to non-essential personnel in the starting, takeoff, and/or flight areas; loose mechanical connections; failed surface and/or range checks; and other forms of mechanical failure.  The safety observer has the absolute authority to give the order to kill the ignition and abort startup immediately upon observing any unsafe condition.  The safety observer is responsible for communicating possible conditions to the startup team as they arise. At minimum, one safety observer is required to start the engine, though we typically have more.

\newpage
\appendix
\renewcommand\thesection{Appendix \Alph{section}}
\section{Preflight Procedures}
\label{sec:appendix_preflight}
\begin{enumerate}
	\setlength{\itemsep}{0cm}
	\setlength{\parskip}{0cm}
	\item Assemble plane.
	\item Check all mechanical and electrical connections.
		\begin{itemize}
		\itemspace
		\item Are all nuts/bolts loc-tited?
		\item Are all servo extensions connected and properly secured?
		\item Are all access hatches secured?
		\end{itemize}
	\item Position wingmen\footnote{Two men, one positioned at each wing tip, preventing the plane from moving during engine start and subsequent surface checks.}, centerman\footnote{Man straddling the fuse, preventing the plane from moving during engine start and subsequent surface checks}.
	\item Pilot turns on transmitter, announces ``transmitter on.''
	\item Pilot requests ``receiver on''.  Centerman turns on receiver, acknowledges.
	\item Pilot completes surface checks:
		\begin{itemize}
		\itemspace
		\item Ailerons
		\item Elevator
		\item Rudder
		\item Throttle
		\end{itemize}
	\item Starter primes engine.
	\item Pilot obtains visual and verbal confirmation that wingmen, centerman are in place and alert.
	\item Pilot requests ``ignition on.''  Centerman moves ignition switch to the ``on'' position and acknowledges.
	\item Starter requests ``prop clear.''  Starter obtains visual confirmation that no person is standing in the prop arc and announces ``prop clear.''
	\item Starter announces ``contact'' and starts the engine.
	\item Starter clears the area.
	\item Pilot repeats surface and throttle checks at close range.
	\item Pilot leaves starting area and performs surface and throttle checks from a distance (range test).
	\item Pilot returns to starting area to prepare for taxi.
	\item Pilot requests ``centerman off.''  Centerman carefully clears the area and acknowledges.
	\item Pilot requests ``wingmen off.''  Wingmen carefully clear the area and acknowledge.
	\item Pilot taxis airplane to appropriate location to begin takeoff.
	\item Pilot commences takeoff.

>>>>

\documentclass[10pt]{report}
\usepackage[pdftex]{graphicx}
\usepackage{hyperref}
\hypersetup{ colorlinks=true
           , linkcolor=blue
           , pdfnewwindow=true 
           }
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{listings}
\usepackage{float}
\usepackage{textcomp}
\usepackage{multicol}
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
\newcommand{\itemspace}{	\setlength{\itemsep}{0cm} \setlength{\parskip}{0cm}}
\newcommand{\degrees}[1]
{
\begin{math}
#1^{\circ} 
\end{math}
}
\hyphenation{auto-nomous construction}

@@

We are advised by
 Professor Tobias Rossmann,
 and several members of the Tri-County R/C Club \cite{tricountyRC}:
 Larry Kosinar,
 Bill Perkowski,
 Mike Korb,
 (the man who donated daedalus), and
 Chris (test pilot) amongst many who spent time mentoring us.

@@

We would like to acknowledge
Jeff Perkins, for his prolific networking and promotion of the team;
Gregg (last name) of BP Hobbies for getting us out of numerous binds;
Dave Devries and Michael Maia, of Invensense, for their technical guidance;
Professor Chung-chieh Shan, for overseeing Pat's independent study on the IMU design;
Ivan Seskar, for keeping us (and many other undergrads) in the black;
Roger Kondos and Brij Pathak, for contributing to the thermopile attitude sensor.

\section{Systems Engineering}

\subsection{Goals}
For this entry, we took a close look at our goals and requirements, and dialed back our approach to ensure we would meet the most basic competition requirements successfully.
We started the design by forming, as a team, some broad goals based on our experience in the 2009 competition.
These formed the basis on which sub-teams and individuals tackled the smaller problems while still keeping an eye on the big picture.

We also wanted to balance building with buying, keeping in mind our slim finances. While its impossible to truly start from scratch\footnote{``If you want to make an apple pie from scratch, you must first create the universe'' - Dr. Carl Sagan \cite{saganquote}}, we tried to build whatever we could given the 16th June competition deadline.

% footnote is broken (breaks from bottom of page one onto bottom of page two), something about the documentclass needs to be changed so it doesn't assume pages are next to each other (book-like)?

@@

Our team was divided by area of expertise into a mechanical team and an electrical team.
The mechanical team began the year with the goal of producing two competition worthy airframes. These were to become the \emph{Daedalus} and \emph{Knight Two}.

We determined, after consulting with R/C experts, that a high wing trainer is the most stable and proven fixed-wing design. Trainers traditionally have light wing loading for slow speed flight, and will be able to carry payload

One of the most difficult design decisions was to quantify ``ample payload.'' The payload size and weight is the major determining factor in the aircraft size.
% -- amanda will finish this by mon noon: 
We started off assuming an electronics (computer + batteries) payload of 2.5lbs and a pan-tilt camera of 2.5lbs and 8 inch diameter.

@@

We felt the Paparazzi system was full featured and proven, and we wanted to adapt it to our needs.

@@

\subsubsection{Daedalus}

A member of our local R/C club, Tri-County RC \cite{tricountyRC}, donated a ten-foot wingspan high wing trainer to our club. The plane, first built out of foam and balsa from long-lost plans in 1986, was accepted graciously, but required quite a bit of work before it was once again flight-worthy. We removed the covering to find water damage and rot which resulted from years of storage. We stripped and rebuilt nearly the entire airframe, adding carbon-fiber reinforcements to the wing in anticipation of increased wing loading due to our payload. Once rebuilt and covered, we tested the \emph{Daedalus} weighed 19.5 lbs with engine. A full 24oz fuel tank, plus 3.5 lbs electronics payload, would bring the maximum takeoff weight to 25.5lbs.

@@

The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.  The \emph{Daedalus} wings supported 64 lbs (250\% of the maximum take-off weight) under static loading.  We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight. Our post-test structure inspection found no warped members or cracked joints.

@@

Figure goes here: Photos of Daedalus in construction, sandbag testing

Figure may go here: Line drawing of Daedalus. JOHN

@@

\begin{figure}
	\centering
	\subfloat[The unfinished \emph{Knight Two}]{\label{fig:knighttwo_construction} 
			\includegraphics[width=0.4\textwidth]{../images/knighttwo_underconstruction.jpg}}
	\subfloat[A completed 12 Foot Telemaster. Photo Credit: Hobby Lobby] {\label{fig:knighttwo_12ft_telemaster} 
			\includegraphics[width=0.4\textwidth]{../images/telemaster_12ft_hobbylobby.jpg}}
  	\label{fig:knighttwo}
  	\caption[Knight Two]{Knight Two}
\end{figure}
Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations

@@

We took advantage of the MHS-PM5's swiveling lens assembly 
(see Figure \ref{fig:camera_frontview}, and added a servo actuator which could rotate the lens \degrees{135} back from straight forward, as shown in Figure \ref{fig:pantilt_linkdetail}. The tilt servo is a Hitech HS 77BB \ref{hitechs77bb}, and is capable of \degrees{180} rotation. Figures \ref{fig:pantilt} and \ref{fig:pantilt_linkdetail} show the camera lens tilted at \degrees{45} below the horizon.
The camera and tilt servo assembley is attached to the aircraft on a rotating mount. A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembly. The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{1260} rotation (3.5 revolutions). Both servos are powered and controlled by the Camera Control board.

@@

\newpage
\appendix
\renewcommand\thesection{Appendix \Alph{section}}
\section{Preflight Procedures}
\label{sec:appendix_preflight}
\begin{enumerate}
	\setlength{\itemsep}{0cm}
	\setlength{\parskip}{0cm}
	\item Assemble plane.
	\item Check all mechanical and electrical connections.
		\begin{itemize}
		\itemspace
		\item Are all nuts/bolts loc-tited?
		\item Are all servo extensions connected and properly secured?
		\item Are all access hatches secured?
		\end{itemize}
	\item Position wingmen\footnotemark, centerman\footnotemark.
	\item Pilot turns on transmitter, announces ``transmitter on.''
	\item Pilot requests ``receiver on''.  Centerman turns on receiver, acknowledges.
	\item Pilot completes surface checks:
		\begin{itemize}
		\itemspace
		\item Ailerons
		\item Elevator
		\item Rudder
		\item Throttle
		\end{itemize}
	\item Starter primes engine.
	\item Pilot obtains visual and verbal confirmation that wingmen, centerman are in place and alert.
	\item Pilot requests ``ignition on.''  Centerman moves ignition switch to the ``on'' position and acknowledges.
	\item Starter requests ``prop clear.''  Starter obtains visual confirmation that no person is standing in the prop arc and announces ``prop clear.''
	\item Starter announces ``contact'' and starts the engine.
	\item Starter clears the area.
	\item Pilot repeats surface and throttle checks at close range.
	\item Pilot leaves starting area and performs surface and throttle checks from a distance (range test).
	\item Pilot returns to starting area to prepare for taxi.
	\item Pilot requests ``centerman off.''  Centerman carefully clears the area and acknowledges.
	\item Pilot requests ``wingmen off.''  Wingmen carefully clear the area and acknowledge.
	\item Pilot taxis airplane to appropriate location to begin takeoff.
	\item Pilot commences takeoff.
